Towards human SuperEEG

### 2018-10-11 paper/main.tex
checking in minor edits

\begin{abstract}
  Human \textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by
    Robert J. Sawyer in his popular science fiction novel \textit{The
      Terminal Experiment}~\cite{Sawy95}} entails measuring ongoing
  neural activity with perfect precision and at arbitrarily high
  spatiotemporal resolution.  Although true SuperEEG is impossible
  using existing methods, here we present a model-based method for
  \textit{inferring} neural activity at millimeter-scale spatial
  resolutions and millisecond-scale temporal resolutions using
  standard human intracranial recordings.  Our approach assumes that
  different people's brains exhibit similar spatial correlations, and
  that (all else being equal) neural activity at nearby locations will
  tend to be similar.  One can then ask, for an arbitrary individual's
  brain: given recordings from a limited set of locations in that
  individual's brain, along with the observed spatial correlations
  learned from other people's recordings, how much can be inferred
  about ongoing activity at \textit{other} locations in that
  individual's brain?\\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

\section*{Introduction}
Modern human brain recording techniques are fraught with
compromise~\citep{SejnEtal14}.  Commonly used approaches include
functional magnetic resonance imaging (fMRI), scalp
electroencephalography (EEG), and magnetoencephalography (MEG).  For
each of these techniques, neuroscientists and electrophysiologists
must choose to optimize spatial resolution at the cost of temporal
resolution (e.g., as in fMRI) or temporal resolution at the cost of
spatial resolution (e.g., as in EEG and MEG).  A less widely used
approach (due to requiring work with neurosurgical patients) is to
record from electrodes implanted directly onto the cortical surface
(electrocorticography; ECoG) or into deep brain structures
(intracranial EEG; iEEG).  However, these intracranial approaches also
require compromise: the high spatiotemporal resolutions of
intracranial recordings comes at the cost of substantially reduced
brain coverage, since safety considerations limit the number of
electrodes one may implant in a given patient's brain.  Further, the
locations of implanted electrodes are determined by clinical, rather
than research, needs.

@@

The current ``gold standard'' for precisely localizing signals and
sampling at high temporal resolution is to take (ECoG or iEEG)
recordings from implanted electrodes (but from a limited set of
locations in any given brain).  This begs the following question: what
can we infer about the activity exhibited by the rest of a person's
brain, given what we learn from the limited intracranial recordings we
have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, which we
call \textit{SuperEEG}, based on Gaussian process
regression~\cite{Rasm06}.  SuperEEG entails using data from multiple
people to estimate activity patterns at arbitrary locations in each
person's brain (i.e., independent of their electrode placements).  We
test SuperEEG approach using two large datasets of intracranial
recordings~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12, EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  We show
that the SuperEEG algorithm recovers signals well from electrodes that
were held out of the training dataset.  We also examine the factors
that influence how accurately activity may be estimated (recovered),
which may have important implications for electrode design and
placement in neurosurgical applications.

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from {0.2}--{3.0}
(total recording time: {0.3}--14.2 hours; Fig.~\suppstats A), and includes data from a list-learning memory experiment
for up to one hour varying by patient with additional buffer time in
which the patients performed typical bed-ridden hospital
activities.  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording
session, ignoring the particular activities or tasks the patients were
performing at any given moment.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might
be able to generate task-general predictions.  Dataset 2 comprised
multi-session recordings from 4436 electrodes implanted in the brains
of 40 epilepsy patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17,
  KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
  KuceEtal18}.  Each recording session lasted from {0.4}--2.2 hours
(total recording time: {0.4}--6.6 hours; Fig.~\suppstats B).  Whereas Dataset 1 included recordings taken during a
wide variety of behaviors, Dataset 2 included recordings taken as each
patient performed each of two memory tasks: a random word list free
recall task and a categorized word list free recall task.  We used
Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional details on both datasets.

>>>>

\begin{abstract}
  Human \textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by
    Robert J. Sawyer in his popular science fiction novel \textit{The
      Terminal Experiment}~\cite{Sawy95}} entails measuring ongoing
  neural activity throughout the entire living brain with perfect
  precision and arbitrarily high spatiotemporal resolution.  Although
  true SuperEEG is impossible using existing methods, here we present
  a model-based method for \textit{inferring} full-brain neural
  activity at millimeter-scale spatial resolutions and
  millisecond-scale temporal resolutions using standard human
  intracranial recordings.  Our approach assumes that different
  people's brains exhibit similar spatial correlations, and that
  activity and correlation patterns are spatially smooth.  One can
  then ask, for an arbitrary individual's brain: given recordings from
  a limited set of locations in that individual's brain, along with
  the observed spatial correlations learned from other people's
  recordings, how much can be inferred about ongoing activity at
  \textit{other} locations in that
  individual's brain?\\\\
  \footnotesize{\textbf{Keywords: Electrocorticography (ECoG),
      intracranial electroencephalography (iEEG), local field
      potential (LFP), epilepsy, maximum likelihood estimation,
      Gaussian process regression}}
\end{abstract}

\section*{Introduction}
Modern human brain recording techniques are fraught with
compromise~\citep{SejnEtal14}.  Commonly used approaches include
functional magnetic resonance imaging (fMRI), scalp
electroencephalography (EEG), and magnetoencephalography (MEG).  For
each of these techniques, neuroscientists and electrophysiologists
must choose to optimize spatial resolution at the cost of temporal
resolution (e.g., as in fMRI) or temporal resolution at the cost of
spatial resolution (e.g., as in EEG and MEG).  A less widely used
approach (due to requiring work with neurosurgical patients) is to
record from electrodes implanted directly onto the cortical surface
(electrocorticography; ECoG) or into deep brain structures
(intracranial EEG; iEEG).  However, these intracranial approaches also
require compromise: the high spatiotemporal resolution of
intracranial recordings comes at the cost of substantially reduced
brain coverage, since safety considerations limit the number of
electrodes one may implant in a given patient's brain.  Further, the
locations of implanted electrodes are determined by clinical, rather
than research, needs.

@@

The current ``gold standard'' for precisely localizing signals and
sampling at high temporal resolution is to take (ECoG or iEEG)
recordings from implanted electrodes (but from a limited set of
locations in any given brain).  This begs the following question: what
can we infer about the activity exhibited by the rest of a person's
brain, given what we learn from the limited intracranial recordings we
have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, which we
call \textit{SuperEEG}, based on Gaussian process
regression~\cite{Rasm06}.  SuperEEG entails using data from multiple
people to estimate activity patterns at arbitrary locations in each
person's brain (i.e., independent of their electrode placements).  We
test our SuperEEG approach using two large datasets of intracranial
recordings~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12, EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  We show
that the SuperEEG algorithm recovers signals well from electrodes that
were held out of the training dataset.  We also examine the factors
that influence how accurately activity may be estimated (recovered),
which may have implications for electrode design and
placement in neurosurgical applications.

\section*{Approach}
The SuperEEG approach to inferring high temporal resolution full-brain
activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate
(in \textit{Results}) our approach using a two large previously
collected dataset comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876
electrodes implanted in the brains of 88 epilepsy
patients~\cite{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
  MannEtal12}.  Each recording session lasted from 0.2--3~h (total
recording time: 0.3--14.2~h; Fig.~\suppstats A).  During each
recording session, the patients participated in a free recall list
learning task, which lasted for up to approximately 1~h.  In addition,
the recordings included ``buffer'' time (the length varied by patient)
before and after each experimental session, during which the patients
went about their regular hospital activities (confined to their
hospital room, and primarily in bed).  These additional activities
included interactions with medical staff and family, watching
television, reading, and other similar activities.  For the purposes
of the Dataset 1 analyses presented here, we aggregated all data
across each recording session, including recordings taken during the
main experimental task as well as during non-experimental time.  We
used Dataset 1 to develop our main SuperEEG approach, and to examine
the extent to which SuperEEG might be able to generate task-general
predictions.  Dataset 2 comprised multi-session recordings from 4436
electrodes implanted in the brains of 40 epilepsy
patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17,
  LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  Each
recording session lasted from 0.4--2.2~h (total recording time:
0.4--6.6~h; Fig.~\suppstats B).  Whereas Dataset 1 included recordings
taken as the patients participated in a variety of activities, Dataset
2 included recordings taken as each patient performed each of two
specific experimental memory tasks: a random word list free recall
task (Experiment 1) and a categorized word list free recall task
(Experiment 2).  We used Dataset 2
to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides
additional information about both datasets.
